Ecosystem size and complexity as extrinsic drivers of food chain length in branching ecosystems

### 2023-03-06 rmd/manuscript.Rmd
Update manuscript.Rmd

## Underlying state changes

@@

The most important finding in this study is the synergistic effects between ecosystem geometry and disturbance. Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain scenario). However, as the frequency of intense disturbance increases, small and/or simple networks could not maintain long food chains because habitats were too homogeneous with limited disturbance refugia. While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction. For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled primarily by the number of tributary confluences in Idaho streams. Similarly, it has been shown that the merging of major tributaries lengthened FCL in the Colorado River, especially during a post-monsoon turbid water phase [@saboPulsedFlowsTributary2018]. More empirical research is needed to confirm our theoretical predictions.

@@

Another interesting result of our simulation is the interaction with omnivory. In particular, when the chain motif was combined with no-disturbance, ecosystem size and complexity had contrasting influences on FCL. It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL. This effect of ecosystem size is understandable because it simply increases the total carrying capacity. In contrast, ecosystem complexity, here assessed as branching probability, decreased the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities. Thus, ecosystem complexity had an opposite effect on FCL with the chain motif. However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios. This fact explains the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.

>>>>

xxx

@@

The most significant finding in this study is the synergistic effects between ecosystem geometry and disturbance. Our no disturbance scenario can be viewed as habitat networks having nearly homogeneous environments, in which long food chains were generally supported independently of ecosystem size and complexity (except the chain scenario). However, as the frequency of intense disturbance increases, small and/or simple networks could not maintain long food chains because habitats were too homogeneous with limited disturbance refugia. While empirical studies rarely looked at the influence of ecosystem complexity on FCL, several studies provide some support for this theoretical prediction. For example, @sullivanEcosystemStructureEmerges2015 reported FCL is controlled primarily by the number of tributary confluences in Idaho streams. Similarly, it has been shown that the merging of major tributaries lengthened FCL in the Colorado River, especially during a post-monsoon turbid water phase [@saboPulsedFlowsTributary2018]. More empirical research is needed to confirm our theoretical predictions.

@@

Another interesting result of our simulation is the interaction with omnivory. In particular, when the chain motif was combined with no-disturbance, ecosystem size and complexity had contrasting influences on FCL. It is evident from the one-patch simulation that the chain motif is highly sensitive to local carrying capacity; therefore, the total carrying capacity within a network dictates FCL. This effect of ecosystem size is understandable because it increases the total carrying capacity. In contrast, ecosystem complexity, here assessed as branching probability, decreased the total carrying capacity as more branched networks have disproportionate numbers of headwater habitats with small local carrying capacities. Thus, ecosystem complexity had an opposite effect on FCL with the chain motif. However, the dependence of FCL on local carrying capacity diminished in weak and strong omnivory scenarios. This fact explains the lack of positive (ecosystem size) or negative FCL-geomtery relationship (ecosystem complexity) in these motifs with no disturbance.
